---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
```{r echo=TRUE}
setwd("E:/myGitFolder/m5_Coursera/RepData_PeerAssessment1")
activity <- read.csv(file = 'activity.csv' , sep = ',')
```
## What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day

```{r echo=TRUE , warning=FALSE}
activityWithoutNA <- subset(x = activity, !is.na(activity$steps))
totalNoOfStepsByDate <- aggregate(x =activityWithoutNA$steps, by = list(activityWithoutNA$date), FUN = sum)
names(totalNoOfStepsByDate)<- c('Activity_Date' , 'Total_Step')

```
Summary : 
```{r}
totalNoOfStepsByDate
```


2. Histogram of the total number of steps taken each day

```{r echo=TRUE , warning=FALSE, message=FALSE}
library(ggplot2)
ggplot(totalNoOfStepsByDate, aes(Total_Step)) + geom_histogram(aes(fill = ..count..)) + labs(title= 'Total number of steps taken each day' , x = 'Steps' , y = "Count" ) 
```

3. Mean and median of the total number of steps taken per day

```{r echo=TRUE , warning=FALSE}
meanTotalStep <-  mean(totalNoOfStepsByDate$Total_Step)
medianTotalStep <-  median(totalNoOfStepsByDate$Total_Step)
```

Mean of the total number of steps taken per day : `r meanTotalStep`

Median of the total number of steps taken per day : `r medianTotalStep`

## What is the average daily activity pattern?



## Imputing missing values


## Are there differences in activity patterns between weekdays and weekends?
